---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.


## Things I Learned on Jan. 27

The command to set a working directory is setwd(). For example:

```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

```{r}
install.packages("tidyverse")
library(tidyverse)
```

### group by and count:

data %>% group_by(COLUMN NAME) %>% summarize(VARIABLE NAME = AGGREGATE FUNCTION(COLUMN NAME))

maryland_winred_contributions %>%
  group_by(memo_text) %>%
  summarise(
    count_contribs = n()
  )
  
### case when

maryland_winred_categories <- maryland_winred %>%
  mutate(
    amount_category = case_when(
        amount < 100 ~ "Less than $100",
        amount >= 100 & amount <= 500 ~ "Between $100 and $500",
        amount > 500 & amount <= 2900 ~ "Between $501 and $2900",
        amount > 2900 ~ "More than $2,900"
      )
  )

### Summarizing

I need to use group_by and summariz. Here's an example of grouping by county and calculating counts, sum and other descriptive statistics.


```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount),
    mean_loan_amount = mean(amount),
    median_loan_amount = median(amount),
    min_loan_amount = min(amount),
    max_loan_amount = max(amount)
  ) %>%
  arrange(desc(max_loan_amount))
```

When you have a new variable:

```{r}
my_variable <- my_new_variable
```

Example:

```{r}
nake<-Ryan
```

You can always save a new codeblcok to a variable. 

###Mutate
Mutate something to add columnes, rows,BUT YOU MUST MUST SAVE IT TO A VARIABLE IN ORDER TO MAKE IT PERMANENT. 

###Interesting Packages:

Lubridate: It allows easier categorization of dates, also allows functions for times. It expands mathematical operations, and 

### When you want to group things by certain groups of numbers:

```{r}
maryland_winred_categories <- maryland_winred %>%
  mutate(
    amount_category = case_when(
        amount < 100 ~ "Less than $100",
        amount >= 100 & amount <= 500 ~ "Between $100 and $500",
        amount > 500 & amount <= 2900 ~ "Between $501 and $2900",
        amount > 2900 ~ "More than $2,900"
      )
  )
head(maryland_winred_categories)
```




```{r}
wv_winred_categories <- wv_winred_contribs %>%
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount <= 25 ~ "Between $10 and $25",
        amount > 25 & amount <= 50 ~ "Between $26 and $50",
        amount > 50 & amount <= 100 ~ "Between $51 and $100",
        amount > 101 & amount <= 501 ~ "Between $101 and $500",
        amount >= 501 ~ "More than $501"
      )
  )
head(wv_winred_categories)
wv_winred_categories %>%
  group_by(amount_category) %>%
  summarize(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
wv_winred_categories %>%
  group_by(amount_category) %>%
  summarise(
    count_contribs = n()
  ) %>%
  arrange((count_contribs))
```


```{r}
 wv_winred_categories_names <- wv_winred_categories  %>% filter(amount_category == "Less than $10")

head(wv_winred_categories_names)
```


Q2. Which West Virginia donors made the most contributions of less than $10? You'll need to use a combination of columns to get unique donors.  
A2. A woman by the name of Shelly Rohrig, who spent a considerable amount of money on caidadtes across the board. She sent in many many donations of less than $10. 

```{r}

less_than_10 <- wv_winred_contribs %>% 
  filter(amount < 10) %>% 
  group_by(first_name, last_name, address_one) %>% 
  summarise(number_of_contributions = n()) %>% 
  arrange(desc(number_of_contributions))
