---
title: "Reverse Engineering Project"
author: "Ryan Mercado"
date: "10/6/2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sentence 1: 

What we are looking for:
we are doing research on how to reverse-engineer from Stephen's paragraph:

“We set the pace in fundraising — including raising more money in Maryland and having 73% of donations $100 or less — because this campaign has inspired thousands of people from all walks of life in Maryland and across the country who see the opportunity to make Wes Moore governor of Maryland and change the trajectories of families who have been left behind,” the campaign said in a statement."

So basically we are looking at contributions of $100 or less. 
This is the process that we used to isolate our data to that point:

Analysis: The data shows that Wes Moore got 8,647 unique contributions of less than 100$


## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)

```

We found Moore contribution data online, created it into a CSV file, and then put it in github to load it: 

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

moore_contribs <- read.csv("data/ContributionsListMoore.csv")
perez_contribs <- read.csv("data/Perez_Contributions.csv")

```



# Clean required data and prepare for analysis if needed. 
```{r}
moore_contribs_cleanednames <- moore_contribs %>% 
  clean_names()
```

```{r}
perez_contribs_cleanednames <- perez_contribs %>% 
  clean_names()

```

```{r}
head(moore_contribs_cleanednames)

```

We Cleaned date: 
```{r}
moore_contribs_edit <-moore_contribs_cleanednames %>% mutate(contribution_date=mdy(contribution_date))

head(moore_contribs_edit)

```

```{r}
perez_contribs_edit <-perez_contribs_cleanednames %>% mutate(contribution_date=mdy(contribution_date))

head(perez_contribs_edit)


```



We set the contribution_amount as a number amount: 
```{r}
perez_contribs_edit_numeric <-perez_contribs_edit %>% 
mutate(across(contribution_amount, as.numeric))


```

```{r}
moore_contribs_edit_numeric <-moore_contribs_edit %>% 
mutate(across(contribution_amount, as.numeric))

```


Then filtered to see contributions under 100:

```{r}
less_than_hundred_moore_contribs <- moore_contribs_edit_numeric %>%
  filter(contribution_amount < 100)

```


```{r}
less_than_hundred_moore_contribs <- less_than_hundred_moore_contribs %>%
  group_by(contributor_name) %>%
  summarise(total = n(), amount = sum(contribution_amount)) %>%
  arrange(desc(total))
less_than_hundred_moore_contribs
```

Analysis: The data shows that Wes Moore got $8,647 unique contributions of less than 100$. It looks like a woman named Delores Parron-Ragland donanted over 82 times to his campaign, a really large amount totaling $1019.30. She most likely is an outlier though as most top donors did not donate anywhere near 80 times. 

### Sentences 2 and 3:

“Moore, a U.S. Army veteran and bestselling author, built a campaign war chest that included over $2.3 million in out-of-state contributions. Tom Perez, a former leader of the Democratic National Committee, raised nearly $1.7 million in donations from outside Maryland.” 

“Moore, a U.S. Army veteran and bestselling author, built a campaign war chest that included over $2.3 million in out-of-state contributions. 





```{r}
md_moore_contribs_states <- moore_contribs_edit_numeric %>%
  filter(str_detect(contributor_address, " MD "))
```

```{r}
moore_contributions_filtered <- moore_contribs_edit_numeric %>%
    mutate(contributor_state = case_when(
    str_detect(contributor_address, " AL ") ~ "Alabama",
    str_detect(contributor_address, " AK ") ~  "Alaska",
    str_detect(contributor_address, " AZ ") ~  "Arizona",
    str_detect(contributor_address, " AR ") ~  "Arkansas",
    str_detect(contributor_address, "5 Regent Dr Queanbeyan AS") ~  "American Samoa",
    str_detect(contributor_address, " CA ") ~  "California",
    str_detect(contributor_address, " CO ") ~ "Colorado",
    str_detect(contributor_address, " CT ") ~  "Connecticut",
    str_detect(contributor_address, " DE ") ~  "Delaware",
    str_detect(contributor_address, " FL ") ~  "Florida",
    str_detect(contributor_address, " GA ") ~  "Georgia",
    str_detect(contributor_address, " HI ") ~ "Hawaii",
    str_detect(contributor_address, " ID ") ~  "Idaho",
    str_detect(contributor_address, " IL ") ~  "Illinois",
    str_detect(contributor_address, " IN ") ~  "Indiana",
    str_detect(contributor_address, " IA ") ~  "Iowa",
    str_detect(contributor_address, " KS ") ~ "Kansas",
    str_detect(contributor_address, " KY ") ~  "Kentucky",
    str_detect(contributor_address, " LA ") ~  "Louisiana",
    str_detect(contributor_address, " ME ") ~  "Maine",
    str_detect(contributor_address, " MD ") ~ "Maryland",
    str_detect(contributor_address, " MA ") ~  "Massachusetts",
    str_detect(contributor_address, " MI ") ~  "Michigan",
    str_detect(contributor_address, " MN ") ~ "Minnesota",
    str_detect(contributor_address, " MS ") ~  "Mississippi",
    str_detect(contributor_address, " MO ") ~  "Missouri",
    str_detect(contributor_address, " MN ") ~  "Montana",
    str_detect(contributor_address, " MT ") ~  "Montana",
    str_detect(contributor_address, " NE ") ~  "Nebraska",
    str_detect(contributor_address, " NV ") ~ "Nevada",
    str_detect(contributor_address, " NH ") ~ "New Hampshire",
    str_detect(contributor_address, " NJ ") ~ "New Jersey",
    str_detect(contributor_address, " NM ") ~ "New Mexico",
    str_detect(contributor_address, " NY ") ~  "New York",
    str_detect(contributor_address, " NC ") ~ "North Carolina",
    str_detect(contributor_address, " ND ") ~ "North Dakota",
    str_detect(contributor_address, " OH ") ~ "Ohio",
    str_detect(contributor_address, " OK ") ~ "Oklahoma", 
    str_detect(contributor_address, " OR ") ~ "Oregon", 
    str_detect(contributor_address, " PA ") ~ "Pennsylvania",
    str_detect(contributor_address, " RI ") ~ "Rhode Island",
    str_detect(contributor_address, " SC ") ~ "South Carolina",
    str_detect(contributor_address, " SD ") ~ "South Dakota",
    str_detect(contributor_address, " TN ") ~ "Tennessee",
    str_detect(contributor_address, " TX ") ~ "Texas",
    str_detect(contributor_address, " UT ") ~ "Utah",
    str_detect(contributor_address, " VT ") ~ "Vermont",
    str_detect(contributor_address, " VA ") ~ "Virginia",
    str_detect(contributor_address, " WA ") ~ "Washington",
    str_detect(contributor_address, " WV ") ~ "West Virginia",
    str_detect(contributor_address, " WI ") ~ "Wisconsin",
    str_detect(contributor_address, " WY ") ~ "Wyoming",
    str_detect(contributor_address, " DC ") ~ "Washington DC",
    TRUE ~ contributor_address
  ))

moore_contributions_filtered
```


```{r}
perez_contributions_filtered <- perez_contribs_edit_numeric %>%
    mutate(contributor_state = case_when(
    str_detect(contributor_address, " MD ") ~ "Maryland",
    str_detect(contributor_address, " NY ") ~ "New York",
    str_detect(contributor_address, " AL ") ~ "Alabama",
    str_detect(contributor_address, " AK ") ~ "Alaska",
    str_detect(contributor_address, " AZ ") ~ "Arizona",
    str_detect(contributor_address, " AR ") ~ "Arkansas", 
    str_detect(contributor_address, " CA" ) ~ "California", 
    str_detect(contributor_address, " CO" ) ~ "Colorado",
    str_detect(contributor_address, " CT ") ~ "Connecticut",
    str_detect(contributor_address, " DE ") ~ "Delaware",
    str_detect(contributor_address, " FL ") ~ "Florida",
    str_detect(contributor_address, " GA ") ~ "Georgia",
    str_detect(contributor_address, " HI ") ~ "Hawaii",
    str_detect(contributor_address, " ID ") ~ "Idaho",
    str_detect(contributor_address, " IL ") ~ "Illinois",
    str_detect(contributor_address, " IN ") ~ "Indiana",
    str_detect(contributor_address, " IA ") ~ "Iowa",
    str_detect(contributor_address, " KS ") ~ "Kansas",
    str_detect(contributor_address, " KY ") ~ "Kentucky",
    str_detect(contributor_address, " LA ") ~ "Louisiana",
    str_detect(contributor_address, " ME ") ~ "Maine",
    str_detect(contributor_address, " MA ") ~ "Massachusetts",
    str_detect(contributor_address, " MI ") ~ "Michigan",
    str_detect(contributor_address, " MN ") ~ "Minnesota",
    str_detect(contributor_address, " MS ") ~ "Mississippi",
    str_detect(contributor_address, " MO ") ~ "Missouri",
    str_detect(contributor_address, " MT ") ~ "Montana",
    str_detect(contributor_address, " NE" ) ~ "Nebraska",
    str_detect(contributor_address, " NV" ) ~ "Nevada ", 
    str_detect(contributor_address, " NJ ") ~ "New Jersey",
    str_detect(contributor_address, " NM ") ~ "New Mexico",
    str_detect(contributor_address, " NY ") ~ "New York",
    str_detect(contributor_address, " NC ") ~ "North Carolina",
    str_detect(contributor_address, " ND ") ~ "North Dakota",
    str_detect(contributor_address, " OH ") ~ "Ohio",
    str_detect(contributor_address, " OK ") ~ "Oklahoma",
    str_detect(contributor_address, " OR ") ~ "Oregon",
    str_detect(contributor_address, " PA ") ~ "Pennsylvania",
    str_detect(contributor_address, " RI ") ~ "Rhode Island",
    str_detect(contributor_address, " SC ") ~ "South Carolina",
    str_detect(contributor_address, " SD ") ~ "South Dakota",
    str_detect(contributor_address, " TN ") ~ "Tennessee",
    str_detect(contributor_address, " TX ") ~ "Texas",
    str_detect(contributor_address, " UT ") ~ "Utah",
    str_detect(contributor_address, " VT ") ~ "Vermont",
    str_detect(contributor_address, " VA ") ~ "Virginia",
    str_detect(contributor_address, " WA ") ~ "Washington",
    str_detect(contributor_address, " WV ") ~ "West Virginia",
    str_detect(contributor_address, " WI ") ~ "Wisconsin",
    str_detect(contributor_address, " WY ") ~ "Wyoming",
    str_detect(contributor_address, " DC ") ~ "Washington, D.C.",
    
    TRUE ~ contributor_address))

```



```{r}
other_variable<-moore_contributions_filtered %>% 
  filter(year(contribution_date) == 2021) %>% 
  group_by(contributor_state) %>% 
  summarise(total_state = sum(contribution_amount),
            mean_amount = mean(contribution_amount),
            n=n()) %>% 
  arrange(desc(total_state))

other_variable

```




```{r}
perez_contributions_filtered %>% 
group_by(contributor_state) %>%
summarise(total_state_amount = sum(contribution_amount),
            mean_amount= mean(contribution_amount),
          number_of_contribs=n())%>%
arrange(desc(total_state_amount))
```
```{r}
moore_out_of_state_contributions <- other_variable %>% 
  filter(!str_detect(contributor_state, "Maryland"))

moore_out_of_state_contributions
```

```{r}
moore_out_of_state_contributions %>%
  summarise(total_state_amount = sum(contribution_amount),
          mean_amount= mean(contribution_amount),
          number_of_contribs=n())%>%
arrange(desc(total_state_amount))


```

```{r}
perez_out_of_state_contributions <- perez_contributions_filtered %>% 
  filter(!str_detect(contributor_state, "Maryland"))
```

```{r}
perez_out_of_state_contributions %>%
  summarise(total_state_amount = sum(contribution_amount),
          mean_amount= mean(contribution_amount),
          number_of_contribs=n())%>%
arrange(desc(total_state_amount))


```
Analysis: From what I can see from the data, Moore got some $4.1 million contributions just from out of state from 9,453 people. And from those people, the mean amount was $434, its a lot of money. Perez, on the other hand got about $2.4 million in out of state contributions, not that bad when comparing his known profile for his work under the Obama administration, it makes sense. What's interesting is that the mean amount was larger than Moore's at a whooping $682 among 3589 unique contributions. Its a lot to donate to someone out of state. Perhaps the Democratic primary was more high-stakes than I thought? 

### Sentences 4:

"John B. King, former secretary of Education under President Barack Obama, raised $2.2 million in campaign funding last year."

```{r}
king_contribs <- read.csv("data/king_contribs.csv")

```

```{r}
king_contribs
```
```{r}
king_contribs_cleanednames <- king_contribs %>% clean_names
```


```{r}
king_contribs_cleanednames
```
```{r}
king_contribs_edit <-king_contribs_cleanednames %>% mutate(contribution_date=mdy(contribution_date))


```


```{r}
king_contribs_edit
```

```{r}
king_contribs_edited<-king_contribs_edit %>% 
  select(-office ,-x)
```

```{r}
king_contribs_2021<-king_contribs_edited %>%
filter(year(contribution_date)==2021)
king_contribs_2021  
```





```{r}
king_contribs_2021 %>%
  summarise(total_2021 = sum(contribution_amount))
```
### Sentences 5:

"Moore and King each received $100,000 each from Illinois." 


###Moore Data:
```{r}
moore_contributions_filtered %>%
    filter(contributor_state == "Illinois") %>% 
  group_by(contributor_state) %>% 
  summarise(sum_Ill = sum(contribution_amount))
```
###King Data:

```{r}
king_contribs_edited_states <- king_contribs_edited %>%
    mutate(contributor_state = case_when(
    str_detect(contributor_address, " MD ") ~ "Maryland",
    str_detect(contributor_address, " NY ") ~ "New York",
    str_detect(contributor_address, " AL ") ~ "Alabama",
    str_detect(contributor_address, " AK ") ~ "Alaska",
    str_detect(contributor_address, " AZ ") ~ "Arizona",
    str_detect(contributor_address, " AR ") ~ "Arkansas", 
    str_detect(contributor_address, " CA" ) ~ "California", 
    str_detect(contributor_address, " CO" ) ~ "Colorado",
    str_detect(contributor_address, " CT ") ~ "Connecticut",
    str_detect(contributor_addre ss, " DE ") ~ "Delaware",
    str_detect(contributor_address, " FL ") ~ "Florida",
    str_detect(contributor_address, " GA ") ~ "Georgia",
    str_detect(contributor_address, " HI ") ~ "Hawaii",
    str_detect(contributor_address, " ID ") ~ "Idaho",
    str_detect(contributor_address, " IL ") ~ "Illinois",
    str_detect(contributor_address, " IN ") ~ "Indiana",
    str_detect(contributor_address, " IA ") ~ "Iowa",
    str_detect(contributor_address, " KS ") ~ "Kansas",
    str_detect(contributor_address, " KY ") ~ "Kentucky",
    str_detect(contributor_address, " LA ") ~ "Louisiana",
    str_detect(contributor_address, " ME ") ~ "Maine",
    str_detect(contributor_address, " MA ") ~ "Massachusetts",
    str_detect(contributor_address, " MI ") ~ "Michigan",
    str_detect(contributor_address, " MN ") ~ "Minnesota",
    str_detect(contributor_address, " MS ") ~ "Mississippi",
    str_detect(contributor_address, " MO ") ~ "Missouri",
    str_detect(contributor_address, " MT ") ~ "Montana",
    str_detect(contributor_address, " NE" ) ~ "Nebraska",
    str_detect(contributor_address, " NV" ) ~ "Nevada ", 
    str_detect(contributor_address, " NJ ") ~ "New Jersey",
    str_detect(contributor_address, " NM ") ~ "New Mexico",
    str_detect(contributor_address, " NY ") ~ "New York",
    str_detect(contributor_address, " NC ") ~ "North Carolina",
    str_detect(contributor_address, " ND ") ~ "North Dakota",
    str_detect(contributor_address, " OH ") ~ "Ohio",
    str_detect(contributor_address, " OK ") ~ "Oklahoma",
    str_detect(contributor_address, " OR ") ~ "Oregon",
    str_detect(contributor_address, " PA ") ~ "Pennsylvania",
    str_detect(contributor_address, " RI ") ~ "Rhode Island",
    str_detect(contributor_address, " SC ") ~ "South Carolina",
    str_detect(contributor_address, " SD ") ~ "South Dakota",
    str_detect(contributor_address, " TN ") ~ "Tennessee",
    str_detect(contributor_address, " TX ") ~ "Texas",
    str_detect(contributor_address, " UT ") ~ "Utah",
    str_detect(contributor_address, " VT ") ~ "Vermont",
    str_detect(contributor_address, " VA ") ~ "Virginia",
    str_detect(contributor_address, " WA ") ~ "Washington",
    str_detect(contributor_address, " WV ") ~ "West Virginia",
    str_detect(contributor_address, " WI ") ~ "Wisconsin",
    str_detect(contributor_address, " WY ") ~ "Wyoming",
    str_detect(contributor_address, " DC ") ~ "Washington, D.C.",
    
    TRUE ~ contributor_address))
```




```{r}
king_contribs_edited_states %>% 
filter(contributor_state == "Illinois") %>% 
  group_by(contributor_state) %>% 
  summarise(sum_Ill = sum(contribution_amount))

```






