---
title: "lab_02"
author: "derek willis"
date: "8/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the slice of WinRed contribution data from West Virginia.  

```{r}
wv_winred_contribs <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the contribution amounts into the following options - less than $10, $10-$25, $26-$50, $51_$100, $101-$500, $501 and up - and calculates how many contributions are in each group. How would you describe the distribution of contributions in a sentence?
A1. Interestingly, or not so interestingly, contributiosn between $10 and $25 are the highest ones. this makes sense since people generally don't want to contribute a large amoutn while 10-25 seems like a reasonable amount to some people. Less than $10 is the next highest, which also makes sense since people don't have a lot of disposable income to send to poltiical campaigns. 


```{r}
wv_winred_categories <- wv_winred_contribs %>%
  mutate(
    amount_category = case_when(
        amount < 10 ~ "Less than $10",
        amount >= 10 & amount <= 25 ~ "Between $10 and $25",
        amount > 25 & amount <= 50 ~ "Between $26 and $50",
        amount > 50 & amount <= 100 ~ "Between $51 and $100",
        amount > 101 & amount <= 501 ~ "Between $101 and $500",
        amount >= 501 ~ "More than $501"
      )
  )
head(wv_winred_categories)
wv_winred_categories %>%
  group_by(amount_category) %>%
  summarize(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
wv_winred_categories %>%
  group_by(amount_category) %>%
  summarise(
    count_contribs = n()
  ) %>%
  arrange((count_contribs))
```

Q2. Which West Virginia donors made the most contributions of less than $10? You'll need to use a combination of columns to get unique donors.  
A2. A woman by the name of Shelly Rohrig, who spent a considerable amount of money on caidadtes across the board. She sent in many many donations of less than $10. 

```{r}
 wv_winred_categories_names <- wv_winred_categories  %>% filter(amount_category == "Less than $10")
head(wv_winred_categories_names)
```

Q3. What questions do the results of Q2 raise in your mind? How could you use the data to explore them? No code here, just your questions.
A3. Why specifically that person? What about her or those like her make so that she has disposable income to send multiple donations of less than $10? Did she perhaps donate a lot over time in the small donation amount and not realize that she spent a lot of money? Is she a member of a large corporation? Is she donating from money from other people? 

Q4. Which month saw the most contributions?
A4. Unsurprisngly, June had the most contributions, and it makes sense since there most likely was a fund-raising blitz before any primaries in other states. Interestingly, the primary elections in WV was on May 10, 2022, so the contributions from June of that month was to races in OTHER states. Very interesting. 

```{r}
wv_winred_categories_names %>% 
  mutate(month = floor_date(date, "month")) %>%
  group_by(month) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
```


Q5. How many contributions of $2,900 - the individual maximum donation - were there, and who did they go to? How many of the recipients are from West Virginia? Do some googling to find out.
A5. Six people made the max contributions and they're some interesting fellows.First off, they're all men, no woman. Some are in academia, while others are in LLCs, or in political jobs. Four of them are NOT from WV, while the others are. One of them, the last one, Duane Zobrist, is a prominent member of the LDS Church, which doesn't really make a lot of sense since I don't think WV has many Mormons. But the LDS Church does spend a lot on political contributions to Republicans. 

```{r}
wv_winred_contribs %>% 
  filter(amount == 2900) %>%
  group_by(last_name, first_name, amount) %>%
  summarise (
    count_contribs = n()
  ) %>% 
  arrange((count_contribs))
  
```
